### 2025-05-30 docs/paper/glossary.tex
initial commit

\newglossaryentry{oop}{
    name={Object-Oriented Programming},
    first={Object-Oriented Programming (OOP)},
    text={OOP},
    description={
            Object-Oriented Programming is a programming paradigm based on the concept of objects, which can hold a state in the form of fields, and methods, in the form of function with access to the object's state.
        }
}
\newglossaryentry{api}{
    name = {Application Programming Interface},
    first = {Application Programming Interface (API)},
    text = {API},
    plural={APIs},
    firstplural={Application Programming Interfaces (APIs)},
    description = {
            An application programming interface (API) defines a way for two software components to communicate with each other.
        }
}
\newglossaryentry{smt}
{
    name={Satisfiability Modulo Theories},
    first={Satisfiability Modulo Theories (SMT)},
    text={SMT},
    description={
            A decision problem for logical formulas with respect to combinations of background theories
            expressed in classical first-order logic with equality.
        }
}
\newglossaryentry{fol}{
    name={First-order logic},
    first={first-order logic (FOL)},
    text={fol},
    description={
            Expressive formal system that allows the following statements:
            \begin{itemize}
                \item Constants ($x, y$)
                \item Functions ($f(x)$)
                \item Relationships ($x > y$)
                \item Connectives ($\land, \lor, \lnot, \implies, \iff$)
                \item Quantifiers ($\forall, \exists$)
            \end{itemize}
        }
}
\newglossaryentry{smtp}{
    name={SMT problem},
    text={SMT problem},
    plural={SMT problems},
    description={
            Decision problems that are true or false depending on whether a given \gls{fol} formula
            is true or false in respect to a given background theory. \\
            In the context of computer science, the background theories are usually include
            arithmetic, bitvectors, arrays, etc.
        }
}
\newglossaryentry{sat}{
    first={Satisfiability (SAT)},
    name={Satisfiability},
    text={SAT},
    description={
            A formula is satisfiable if it is possible to find a set of values for its variables
            that makes the formula true.
            On the other hand, a formula is unsatisfiable if such set does not exists.
            \begin{multline*}
                \text{Example} \\
                (a \lor c) \land (b \lor c) \land (\neg a \lor \neg c) \\
                \text{is satisfiable with the assignment } a = 1, b = 1, c = 0 \\
            \end{multline*}
        }
}
\newglossaryentry{convopt}{
    name={Convex optimization},
    text={convex optimization},
    description={
            A convex optimization problem is a problem where the objective function and the constraints
            are convex functions. \\
            A convex function is a function whose domain is a convex set and that satisfies the following
            property:
            \begin{equation*}
                f(\lambda x + (1 - \lambda)y) \leq \lambda f(x) + (1 - \lambda)f(y)
            \end{equation*}
            for all $x, y \in \text{dom}(f)$ and $\lambda \in [0, 1]$.
        }
}
\newglossaryentry{cnf}{
    first={Conjunctive Normal Form (CNF)},
    name={Conjunctive Normal Form},
    text={CNF},
    description={
            A formula is in \gls{cnf} if it is a conjunction of clauses, where a clause is a
            disjunction of literals. \\
            A literal can be a boolean variable or its negation. \\
            \\
            \begin{equation*}
                \bigwedge_{i=1}^n \bigvee_{j=1}^{m_i} l_{ij}
            \end{equation*}
        }
}
\newglossaryentry{lp}{
    first={Linear Programming (LP)},
    name={Linear Programming},
    text={LP},
    description={
            A Linear Programming problem is an optimization problem where the objective function and the constraints
            are linear functions. \\
            A typical Linear Programming problem has the following form:
            \begin{equation*}
                \begin{aligned}
                     & \text{maximize}   & c^T x      \\
                     & \text{subject to} & A x \leq b \\
                     &                   & x \geq 0
                \end{aligned}
            \end{equation*}
            where $x \in \mathbb{R}^d$ is the vector of variables to be determined, $c \in \mathbb{R}^d$ and $b \in \mathbb{R}^n$ are vectors of coefficients, and $A \in \mathbb{R}^{n \times d}$ is a matrix of coefficients.
        }
}
\newglossaryentry{qf-lia}{
    name={QF\_LIA},
    text={QF\_LIA},
    description={
            Quantifier-free linear integer arithmetic. In essence, Boolean combinations of inequations between linear polynomials over integer variables.
        }
}
\newglossaryentry{qf-lra}{
    name={QF\_LRA},
    text={QF\_LRA},
    description={
            Quantifier-free linear real arithmetic. In essence, Boolean combinations of inequations between linear polynomials over real variables.
        }
}
\newglossaryentry{nnf}{
    name={Negation Normal Form},
    text={NNF},
    first={Negation Normal Form (NNF)},
    description={
            A formula where the negation operator $\neg$ is only applied to variables. \\
            $\neg a$ is a NNF formula, while $\neg (a \lor b)$ is not.
        }
}
\newglossaryentry{dpll}{
    name={DPLL algorithm},
    text={DPLL},
    first={Davis-Putnam-Logemann-Loveland (DPLL)},
    description={
            A complete \gls{sat} backtracking algorithm for deciding the satisfiability of propositional logic formulae in \gls{cnf}.
        }
}
\newglossaryentry{dp}{
    name={DP algorithm},
    text={DP},
    first={Davis-Putnam (DP)},
    description={
            A complete \gls{sat} algorithms that uses the resolution inference rule to decide the satisfiability of propositional logic formulae in \gls{cnf}.
        }
}
\newglossaryentry{mps}{
    name={Mathematical Programming System},
    text={MPS},
    first={Mathematical Programming System (MPS)},
    description={
            A common file format for representing \gls{lp} problems.
            Developed by IBM, it is supported by most commercial \gls{lp} solvers.
        }
}
\newglossaryentry{soi}{
    name={Sum of Infeasibilities},
    text={SoI},
    first={Sum of Infeasibilities (SoI)},
    description={
            A metric used to evaluate the quality of an \gls{lp} relaxation.
            It is the sum of the infeasibilities of the constraints that are violated by the solution.
        }
}
\newglossaryentry{onnx}{
    name={Open Neural Network Exchange},
    text={ONNX},
    first={Open Neural Network Exchange (ONNX)},
    description={
            An open-source format for representing deep learning models.
            It is supported by a wide range of tools and libraries.
        }
}
\newglossaryentry{cdlc}{
    name={Conflict Driven Learning Clause},
    text={CDLC},
    first={Conflict Driven Learning Clause (CDLC)},
    description={
            A clause that is added to the \gls{cnf} formula after a conflict is detected during the \gls{sat} solving process.
            It is used to prevent the solver from exploring the same path that led to the conflict.
        }
}