### 2025-06-09 docs/paper/main.tex
feat: add tuning

\documentclass{article}

\input{setup.tex}

\begin{document}

\title{Delta-Complete Linear Programming}

\author{Oliver Sch\"{o}n \and
Ernesto Casablanca \and
Sadegh Soudjani \and
Paolo Zuliani
}

\begin{comment}
\institute{O. Sh\"{o}n,E. Casablanca \at Newcastle University, Newcastle upon Tyne, United Kingdgom \\ \email{e.casablanca2@newcastle.ac.uk}, \email{o.schoen2@newcastle.ac.uk}
\and
S. Soudjani \at Max Planck Institute for Software Systems, Kaiserslautern, Germany \\ \email{sadegh@mpi-sws.org}
\and
P. Zuliani \at Universit\`{a} di Roma ``La Sapienza'', Rome, Italy\\ \email{zuliani@di.uniroma1.it}
}
\end{comment}

\maketitle

\begin{abstract}
  Ensuring the safety of AI-enabled systems, particularly in high-stakes domains like autonomous driving and healthcare, has become increasingly critical.
  Traditional formal verification tools fall short when faced with the opaque, black-box nature of AI components and the sheer scale of modern applications.
  To address these challenges, we introduce Lucid, a new tool designed specifically for the safety verification and synthesis of black-box systems governed by discrete-time stochastic dynamics.

  Lucid employs a data-driven methodology rooted in control barrier certificates, which are learned directly from system trajectory data to ensure formal safety guarantees.
  We use conditional mean embeddings to embed data into a reproducing kernel Hilbert space (RKHS), where it constructs an RKHS ambiguity set that can be inflated to robustify the result to out-of-distribution behavior.
  To generalize beyond safety, we provide theoretical foundations for applying Lucid to broader classes of temporal logic specifications.

  A key innovation within Lucid is its use of a finite Fourier expansion to reformulate a semi-infinite non-linear optimization problem into a tractable linear program.
  The resulting spectral barrier allows us to leverage the fast Fourier transform to generate the relaxed problem efficiently, offering a scalable yet distributionally robust framework for verifying safety.
  Lucid thus offers a robust and efficient verification framework, able to handle the complexities of modern black-box systems while providing formal guarantees of safety.
\end{abstract}

\section{Introduction}
\label{sec:intro}

Lucid is in implemented in C++ and it is meant to be used as a library.
It provides a set of interfaces that allow users to have an high level of control and flexibility over the verification process.
We also provide a Python wrapper, called PyLucid, to facilitate the integration of the tool into existing workflows and effortlessly leverage well-established libraries such as NumPy~\cite{numpy} and SciPy~\cite{scipy}.
Lucid is available as an open-source project on GitHub at \url{https://github.com/TendTo/lucid}.
More technical details about the implementation can be found in the documentation at \url{https://tendto.github.io/lucid/}.

\section{Problem Statement}
\label{sec:problem-statement}

\subsection{Notation}
\label{sec:notation}

% Spaces and numbers
We denote the sets of positive integers and non-negative reals as $\N_{>0}$ and $\R_{\geq 0}$, respectively.
% Probability theory
Consider a Polish sample space $\X$ \cite{bogachev2007measure}.
Let $(\X,\borel{\X},\P)$ be the underlying probability space equipped with a Borel $\sigma$-algebra $\borel{\X}$ defined over $\X$, i.e., the smallest $\sigma$-algebra containing open subsets of $\X$, and a probability measure $\P$.
For a random variable $X$, let $p_X$ be the pushforward probability measure of $\P$ under $X$ such that $X\sim p_X(\cdotx)$.

The expected value of a function $f(X)$ on $\X$ is written as $\E_{p_X}[f(X)]$. If it is clear from the context, we abbreviate and write $\E[f(X)]$.
We denote the set of all probability measures for a given measurable space $(\X,\borel{\X})$ as $\mathcal{P}(\X)$.

The $n$-dimensional Gaussian measure with mean $\mu\in\mathbb{R}^n$ and covariance matrix $\Sigma\in\mathbb{R}^{n\times n}$ is given by
\begin{equation}
  \label{eq:gaussian-kernel}
  \mathcal N(dx|\mu, \Sigma) := \frac{dx}{ \sqrt{(2\pi)^n \left|\Sigma\right| }}\exp\left[-\frac{1}{2}(x-\mu)\T\Sigma^{-1\!}(x-\mu)\right],
\end{equation}
where $|\Sigma|$ is the determinant of $\Sigma$.
The Dirac delta measure $\delta_a:\borel{\X}\rightarrow [0,1]$ concentrated at a point $a\in\X$ is defined as $\delta_a(A)=1$ if $a\in A$ and $\delta_a(A)=0$ otherwise, for any measurable set $A\in\borel{\X}$.
We denote the uniform distribution over $\X$ as $\mathcal{U}_\X$ with realizations $x\sim\mathcal{U}_\X(\cdotx)$.

For two measurable spaces $(\X,\borel{\X})$ and $(\Y,\borel{\Y})$, a \emph{probability kernel} is a mapping $\p: \X \times \borel{\Y}\rightarrow  [0,1]$ such that $\p(X=x,\cdotx):\borel{\Y}\rightarrow[0,1]$ is a probability measure for all $x\in\X$, and $\p(\cdotx, B): \X\rightarrow [0,1]$ is measurable for all  $B\in\borel{\Y}$.
A probability kernel associates to each point $x\in\X$ a measure denoted by $\p(\cdotx|X=x)$.

% Linear algebra
The transpose of a vector or matrix $A$ is indicated as $A\T$.
Let the $N\times N$ dimensional identity matrix be given by $I_N$.
Let $X_N:=[x_i]_{i=1}^N$ be a column vector with $x_i\in\X$.
We denote the element-wise evaluation of a function $f:\X\rightarrow\R$ on $X_N$ as $f(X_N):=[f(x_i)]_{i=1}^N$.
Similarly, we may write $A = [a_{ij}]_{i,j=1}^N$ to denote a matrix with its elements.

\subsection{Discrete-Time Stochastic Systems}
In this work, we consider systems expressible as Markov decision processes over continuous state and input spaces, formally defined as follows.
\begin{definition}[Markov decision process (MDP)]
  An MDP is a tuple $\M=(\X,\X_0,\U,\Tr)$, comprising a state space $\X\subset\R^n$ with states $x\in\X$; initial states $x_0\in \X_0\subset\X$; an input space $\U$ with inputs $u\in\U$; and a probability kernel $\Tr:\X\times\U\times\borel{\X}\rightarrow[0,1]$.
\end{definition}

In every execution, given a current state $x\in\X$ and control input $u\in\U$, the MDP evolves to a consecutive state $x_{+}\in\X$, which is obtained as a realization $x_{+}\sim\Tr(\,\cdot\,|X=x,U=u)$.
As a notable class of systems that can be captured using MDPs, we may consider black-box systems with Markovian discrete-time stochastic dynamics, namely
\begin{equation}
  \label{eq:model}
  \M: \left\{ \begin{array}{ll}
    x_{t+1} & = f(x_t,u_t,w_t),\quad w_t\sim p_w(\cdotx), \\
    y_t     & = x_t,
  \end{array} \right.
\end{equation}
where the system state, control input, and observation at the $t^{\text{th}}$ time-step are denoted by $x_t$, $u_t$, and $y_t$, respectively.
The state evolution of the system, described by the function $f:\X\times\U\times\mathbb{W}\rightarrow\X$,
is subject to {independent, identically distributed (i.i.d.) noise} $w_t\sim p_w(\cdotx)$ supported on a set $\mathbb{W}$.

\paragraph{Implementation}

When using \lucid, the user needs to provide a set of sampled trajectories from the system.
This is often achieved by defining a function that takes as input the current state and control input and returns the next state, adding an i.i.d. noise sampled from an appropriate distribution, or it may be collected from a simulator or a real-world system.
The sampled trajectories are then fed into \lucid, either programmatically or through a file \todo{CSV, HDF5, other?}.
\todo{Allow a simple way to define the system model? Maybe latex2sympy?}
\todo{Graphical interface?}

\subsection{System Safety}
In this paper, we wish to synthesize a controller $\pi$ such that the closed-loop system $\pi\times\M$ is safe.
Particularly, we restrict our attention to \emph{stationary} policies of the form $\pi:\X\rightarrow\U$, as for MDPs there always exists a stationary policy that is optimal.
For an MDP $\M$, we characterize \emph{safety} as a property of its state. More precisely, a system $\pi\times\M$ is said to be safe if for any initial state $x_0\in \X_0$ the system does not enter an \emph{unsafe set} of states $\X_u\subset\X$.
Note that general stochastic systems do not admit a binary notion of safety.
Thus, we quantify the \emph{probability} of reaching $\X_u$ in a given finite horizon $T\in\N_{>0}$ and compute its complement.
A (finite-horizon) \emph{safety specification} $\psi_{\text{safe}:=(\X_u,T)}$ is hence fully characterized by an unsafe set $\X_u$ and a time horizon $T$.
The system $\pi\times\M$ satisfies $\psi_{\text{safe}}$ with probability at least $p\in(0,1)$ if the probability of its trajectories starting in $\X_0$ and avoiding $\X_u$ within horizon $T$ is at least $p$. This is denoted by $P^\pi_{\text{safe}}(\M):=\P(\pi\times\M\satisfies\psi_{\text{safe}})\geq p$~\cite{BK08}.

\paragraph{Implementation}

\lucid understands \texttt{RectSet}s and \texttt{Multiset}s \todo{PolytopeSet?}.
\todo{Set definition from file?}.
These are used to indicate the initial $\X_0$ set and unsafe set $\X_u$ when defining the safety specification $\psi_{\text{safe}}$.
Once the system is fully defined, the time horizon $T$ is set as a parameter of the verification process.
\todo{Support other kinds of properties other than safety (e.g., reachability)?}

\subsection{Control Barrier Certificates}
\label{sec:cbc}
Whilst certifying the safety of a stochastic system is generally challenging, \emph{control barrier certificates} (CBCs) and \emph{control barrier functions} (CBFs) leverage the concept of set invariance to arrive at an abstraction-free formulation. This has made CBCs/CBFs popular tools for safety verification and synthesis~\cite{prajna2006barrier,cosner2023generative}.
We briefly recall the definition of a CBC.
\begin{definition}[Control barrier certificate (CBC)]\label{def:cbc}
  A function $\B:\X\rightarrow\R_{\geq 0}$ is called a CBC of an MDP $\M=(\X,\X_0,\U,\Tr)$ with reference to an unsafe set $\X_u$, if we have
  \begin{itemize}
    \item[(a)] $\forall x_0\in \X_0\!:\,\B(x_0)\leq\eta$;
    \item[(b)] $\forall x_u\in \X_u\!:\,\B(x_u)\geq\gamma$; and
    \item[(c)] $\forall x\in\X,\exists u\in\U\!:\,\E_{\Tr}[\,\B(X^+)\mid X=x,\,U=u\,] -\B(x) \leq  c;$
  \end{itemize}
  for some constants $\gamma>\eta\geq0$ and $c\geq 0$.
\end{definition}
If one can find a CBC for a system $\M$, then, a lower bound on the probability of $\M$ being safe is given by the following proposition due to \cite{kushner1967stochastic}.
\begin{proposition}[Finite-horizon safety]
  \label{prop:reach}
  Consider an MDP $\M=(\X,\X_0,\U,\Tr)$ and a safety specification $\psi_{\text{safe}}=(\X_u,T)$. Suppose there exists a CBC $B$ w.r.t. $\X_u$ (Definition~\autoref{def:cbc}) with constants $\eta$, $\gamma$, and $c$. Then, there exists a stationary policy $\pi$ such that
  $$P^\pi_{\text{safe}}(\M)\geq 1- \frac{\eta + cT}{\gamma}.$$
\end{proposition}

\paragraph{Implementation}

When verifying the safety of a system, \lucid will search for a CBC $\B$ that satisfies the conditions in Definition~\ref{def:cbc}.
If it finds it, it will return the constants $\eta$, $\gamma$, and $c$, as well as the CBC $\B$ itself, which becomes a certificate of safety for the system $\M$.
The template of the barrier function is explored later in section~\ref{sec:benchmarks}.

\subsection{Kernel Mean Embeddings}

A symmetric function $k_\X:\X\times\X\rightarrow\R$ is called a (positive definite) \emph{kernel} (note the distinction from \emph{probability kernels}) if for all $N\in\N_{>0}$ we have $\sum_{i=1}^{N}\sum_{j=1}^{N}a_i a_j\allowbreak k_\X(x_i,x_j) \geq 0$ for $x_1,\ldots,x_N\in\X\subset{\R^n}$ and $a_1,\ldots,a_N\in\R$.
A prominent example is the \emph{squared exponential} (SQExp) kernel \cite{Rasmussen2005GP,Kanagawa2018GPvsKernel}:
\begin{equation}
  \label{eq:gaussian-kernel}
  k_\X(x,x') := \sigma_f^2 \exp\left( -\frac{1}{2} (x-x')\T \Sigma\, (x-x') \right),\quad \Sigma:=\diag(\sigma_l)^{-2},
\end{equation}
with amplitude $\sigma_f^2\geq0$ and lengthscale coefficients $\sigma_l\in\R^n$.
%In the following, we refer to a positive definite kernel as \emph{kernel}, for simplicity.
For this work, we assume that all kernels are bounded on their domain, i.e., $\E_{}[k_\X(x,x)]<\infty$, $x\in\X$.
Given a kernel $k_\X$ on a non-empty set $\X$, there exists a unique corresponding \emph{reproducing kernel Hilbert space} (RKHS) $\Hilbert_{k_\X}$ of functions $f:\X\rightarrow\R$ equipped with an inner product $\innerH{\cdotx}{\cdotx}{\Hilbert_{k_\X}}$ possessing the \emph{reproducing property} such that for any function $f\in\Hilbert_{k_\X}$ and $x\in\X$ we have $f(x)=\innerH{f}{k_\X(\cdotx,x)}{\Hilbert_{k_\X}}$.
Note that $k_\X(\cdotx,x):\X\rightarrow\Hilbert_{k_\X}$ is a real-valued function,
\new{which is} also called an implicit \emph{canonical} \emph{embedding} or \emph{feature map} $\phi_\X$
such that $k_\X(x,x')=\innerH{\phi_\X(x)}{\phi_\X(x')}{\Hilbert_{k_\X}}$ for all $x,x'\in\X$.
For an RKHS $\Hilbert_{k_\X}$, we use the associated feature map $\phi_\X$ and kernel $k_\X$ interchangeably for ease of notation and comprehensibility.
%We will call the kernel associated to an RKHS $\Hilbert_{k_\X}$ the \emph{characteristic} kernel of $\Hilbert_{k_\X}$ and use $\phi_\X$ and $k_\X$ interchangeably for ease of notation and comprehensibility.
%and an implicit \emph{canonical} \emph{embedding} or \emph{feature map} $\phi_\X:\X\rightarrow\Hilbert_{k_\X}$ with an inner product on $\Hilbert_{k_\X}$, i.e., $k_\X(x,x')=\innerH{\phi_\X(x)}{\phi_\X(x')}{\Hilbert_{k_\X}}$, 
%Furthermore, 
The inner product induces the norm $\norm{f}_{\Hilbert_{k_\X}}\!\!\!\!:=\!\!\sqrt{\smash[b]{\innerH{f}{f}{\Hilbert_{k_\X}}}}$ of the RKHS.
%Note that $\Hilbert_{k_\X}$ is a space of functions $f:\X\rightarrow\R$. 
%It is equipped with the infamous \emph{reproducing property} such that for any function $f\in\Hilbert_{k_\X}$ and $x\in\X$ we have $f(x)=\innerH{f}{\phi_\X(x)}{\Hilbert_{k_\X}}$.
%Furthermore, the norm associated with the RKHS $\Hilbert_{k_\X}$ is defined as $\norm{f}_{\Hilbert_{k_\X}}:=\sqrt{\innerH{f}{f}{\Hilbert_{k_\X}}}$.
%We refer the interested reader to XX for a comprehensive dissemination on RKHS.
Throughout this paper, we assume that all RKHSs are \emph{separable}.
Given $N$ i.i.d. samples $\hat X_N:=[\hat{x}_i]_{i=1}^N$ with $\hat{x}_i\in\X$, the
\emph{Gram matrix} of $k_\X$ is given by
$K_{\hat{X}}^N:=[k_\X(\hat{x}_i,\hat{x}_j)]_{i,j=1}^N.$
Furthermore, we define the vector-valued function
$k_{\hat{X}}^N(x)  := [k_\X(x,\hat{x}_i)]_{i=1}^N.$

\subsection{Data-Driven Barrier Formulation}
\label{sec:data-driven-barrier-formulation}
Compared to the typical spatial form \eqref{eq:representer-form} based on spatial support vectors $\bar{x}_1,\bar{x}_2,\ldots\in\X$, the spectral form \eqref{eq:representer-form-spectral} is a much more efficient representation for higher-dimensional systems.
The number of spatial support vectors needed to construct adequately expressive barriers grows exponentially in the dimensionality of the system. We avoid this.

Apart from the memory efficiency of the spectral barrier form, we can apply the same idea to collapse the complex CME term in \eqref{eq:semiinf-prog-kushner} via the \emph{fast Fourier transform} (FFT). That is we obtain spectral coefficients $H\in\R^{(2M+1)\times (2M+1)}$ s.t.
\begin{equation}
  k^N_{\hat{X}}(x)\T\left[ K^N_{\hat{X}}+N\lambda I_N\right]^{-1\!} (\Phi_{M, \hat{X}^{+\!}}^{N})\T b \approx \phi_M(x)\T H b,\label{eq:cme-term-spectral-approx}
\end{equation}
where $\Phi_{M, \hat{X}^{+\!}}^{N} := [\phi_M(\hat{x}^+_1),\ldots,\phi_M(\hat{x}^+_N)]$.
Whilst projection onto a truncated Fourier basis introduces an approximation error, the error decreases exponentially with the cardinality of wavenumbers in the basis \cite{rahimi2007random}.
% Our numerical results suggest that the empirical error is negligible.

Substituting the barrier $B\in\Hilbert_M$ of the form \eqref{eq:representer-form-spectral} and the approximation of the CME term in \eqref{eq:cme-term-spectral-approx} into the SIP \eqref{eq:semiinf-prog}, we obtain the \emph{semi-infinite linear program} (SILP)
\begin{subequations}
  \begin{align}
    \min_{b, c, \eta, \gamma} \; & \frac{\eta + cT}{\gamma},\label{eq:semiinfLP-prog-objective}                                                 \\
    \text{s.t.}\quad
                                 & \forall x_0\in\X_0:\,\phi_M(x_0)\T b\leq\eta,\label{eq:semiinfLP-prog-initial}                               \\
                                 & \forall x_u\in\X_u:\,\phi_M(x_u)\T b\geq\gamma,\label{eq:semiinfLP-prog-unsafe}                              \\
                                 & \forall x\in\X: \, \phi_M(x)\T (H b - b) \leq c - \varepsilon\bar{B}\kappa,\label{eq:semiinfLP-prog-kushner} \\
                                 & \forall x\in\X:\,\phi_M(x)\T b\geq 0,\label{eq:semiinfLP-prog-positive}                                      \\
                                 & c\geq 0,\,\gamma>\eta\geq 0,\,b\in\R^{2M+1},\nonumber
  \end{align}\label{eq:semiinfLP-prog}%
\end{subequations}
with coefficients $\kappa\geq\sigma_f$ and $\bar{B}\geq\norm{b}_2$.

\subsection{Finite-Constraint Relaxation}\label{sec:finite-constraint-relaxation}
There exist results that establish deterministic sampling-based bounds for multivariate trigonometric polynomials \cite{cifuentes2017sampling,pfister2018bounding} that can be applied to the bandlimited spectral barrier in \eqref{eq:representer-form-spectral}. This allows us to relax the SILP \eqref{eq:semiinfLP-prog} to a \emph{linear program} (LP) with finitely many constraints by sampling a finite set
$X_{\hat{N}} \subset\X$ of cardinality $\hat{N}$.
By choosing $\hat{N}=Q^n$ equidistant points that form a discrete lattice of points to cover the (state) space, the corresponding constraints can be efficiently evaluated in $\mathcal{O}(Q\log(Q))$ operations using the inverse FFT.
Note that this also substantially reduces the number of evaluations of the computationally expensive CME term in \eqref{eq:semiinf-prog-kushner}, which requires $\mathcal{O}(N^2)$ operations per evaluation. Refer to the Petersen-Middleton theorem \cite{petersen1962sampling}, which generalizes the Nyquist-Shannon sampling theorem to higher-dimensional Euclidean spaces, for further details.

We use the following result by \cite{pfister2018bounding}.
\begin{lemma}[Trigonometric bounds]\label{lem:trig-bound}
  Consider a real-valued trigonometric polynomial $f:[0,1]^n\rightarrow\R$ (cf. \eqref{eq:representer-form-spectral}) with maximum degree $d\in\N_{\geq 0}$ and a sampling lattice $\Theta_Q:=\{\frac{2\pi l}{Q},\, l=0,\ldots,Q-1\}$ with $Q\geq 2d+1$.
  Let $\bar{f}_{Q^n}:=\max_{z\in\Theta^n_Q} f(z)$ and $\underbar{f}_{Q^n}:=\min_{z\in\Theta^n_Q} f(z)$. Then, we have that
  \begin{equation*}
    \frac{1}{2}\big(\bar{f}_{Q^n} + \underbar{f}_{Q^n} - C_{Q^n}\left(\bar{f}_{Q^n}-\underbar{f}_{Q^n}\right)\big)
    \leq
    f(z)
    \leq
    \frac{1}{2}\big(\bar{f}_{Q^n} + \underbar{f}_{Q^n} + C_{Q^n}\left(\bar{f}_{Q^n}-\underbar{f}_{Q^n}\right)\big),
  \end{equation*}
  with a coefficient given by
  \setlength{\belowdisplayskip}{3pt}
  \begin{equation*}
    C_{Q^n} := \left(1-\frac{2d}{Q}\right)^{-\frac{n}{2}}.
  \end{equation*}
\end{lemma}
% \Oliver{The paper \cite{pfister2018bounding} provides a more complex expression for $C_{Q^n}$ that I put as [...]. They also provide the upper bound I use so far.}

Via Lemma~\ref{lem:trig-bound}, we obtain a finite-constraint relaxation of the SIP \eqref{eq:semiinf-prog} based on the constraint-tightening reasoning
\begin{equation*}
  \frac{1}{2}\big(\mathrm{max}_{\hat{N}}^\X + \mathrm{min}_{\hat{N}}^\X - C_{\hat{N}}\left(\mathrm{max}_{\hat{N}}^\X-\mathrm{min}_{\hat{N}}^\X\right)\big)\geq \mathrm{lb}
  \quad\Longrightarrow\quad
  \forall x\in\X: f(x) \geq \mathrm{lb},
\end{equation*}
where $\mathrm{max}_{\hat{N}}^\X:=\max_{x\in X_{\hat{N}}} \left\lbrace\phi_M(x)\T b\right\rbrace$ and $\mathrm{min}_{\hat{N}}^\X:=\min_{x\in X_{\hat{N}}} \left\lbrace\phi_M(x)\T b\right\rbrace$ for some desired lower bound $\mathrm{lb}\in\R$.
Thus, we obtain strengthened constraints of the form
\begin{align*}
                               & \frac{1}{2}\big(\mathrm{max}_{\hat{N}}^\X + \mathrm{min}_{\hat{N}}^\X - C_{\hat{N}}\left(\mathrm{max}_{\hat{N}}^\X-\mathrm{min}_{\hat{N}}^\X\right)\big) \geq \mathrm{lb}, \\
  \Leftrightarrow\hspace{10pt} &
  \mathrm{min}_{\hat{N}}^\X \geq \frac{2}{C_{\hat{N}}+1}\mathrm{lb} + \frac{C_{\hat{N}}-1}{C_{\hat{N}}+1}\mathrm{max}_{\hat{N}}^\X.
\end{align*}
For an upper bound $\forall x\in\X: f(x) \leq \mathrm{ub}$, we obtain analogous constraints of the form
\begin{equation*}
  \mathrm{max}_{\hat{N}}^\X \geq \frac{2}{C_{\hat{N}}+1}\mathrm{ub} + \frac{C_{\hat{N}}-1}{C_{\hat{N}}+1}\mathrm{min}_{\hat{N}}^\X.
\end{equation*}

\subsection{Tuning the Hyperparameters}
\label{sec:tuning-hyperparameters}

Kernel methods do not require the expensive learning process of other machine learning methods, such as neural networks.
However, they still depend on a number of \hp, such as the kernel bandwidth $\sigma_f$ and the number of Fourier coefficients $M$.
Changes in their values can have a significant impact on the estimator's efficiency and accuracy.
The process of finding good values for these \hp is known as \emph{hyperparameter tuning}, and it is problem-dependent.

There are several approaches to find the optimal \hp for a given problem.
A cheap option which can be used as a starting point when working with a Gaussian Kernels as in~\eqref{eq:gaussian-kernel} is to use the \emph{median heuristic} \cite{garreau2018largesampleanalysismedian}.
We choose $\sigma_l^2$ to be the median of the squared distances between all pairs of points in the training set, i.e.,
\begin{equation}
  \sigma_l^2 = \mathrm{median}_{1 \leq i < j \leq N}\left(\|\hat{x}_i-\hat{x}_j\|^2\right),
\end{equation}
where $\hat{x}_1,\ldots,\hat{x}_N$ are the $N$ training samples.
If $N(N-1)/2$ is even, the median is the average of the two middle values in the sorted list of distances.
Better results can be obtained via \emph{grid search}, which consists of evaluating the performance of the estimator for all combinations of \hp in a predefined grid.
This approach is simple to implement and can be very effective for small problems, but it can become computationally expensive for larger problems.
Borrowing a technique from Gaussian processes, we can also try to optimize the \hp to maximize the \emph{log marginal likelihood}, defined as
\begin{equation}
  \log p(\hat{X^+}_N | \hat{X}_N) = - \frac{1}{2}\hat{X^+}_N\T (K_{\hat{X}}^N + N\lambda I_N)^{-1} \hat{X^+}_N -\frac{1}{2}\log |K_{\hat{X}}^N + N\lambda I_N| - \frac{N}{2}\log(2\pi).
\end{equation}

\paragraph{Implementation}

While it is possible to tune the \hp with any method and then set them manually in the estimator, \lucid provides a set of classes, which specialize the \texttt{Tuner} interface, to automate the process.
The \texttt{GridSearchTuner} class implements the grid search method, while the \texttt{MedianHeuristicTuner} class implements the median heuristic.
If the user wants to optimize the log marginal likelihood, they can use the \texttt{LBGSTuner} class, which runs the L-BFGS or L-BFGS-B quasi-Newton optimization algorithms~\cite{book:nonlinear-programming} over the log marginal likelihood function.
Internally, \lucid uses the \texttt{LBFGS++} library~\footnote{https://github.com/yixuan/LBFGSpp} to perform the optimization.

\section{Benchmarks}
\label{sec:benchmarks}

\section*{Acknowledgements}
The authors thank \dots.

\bibliographystyle{ieeetr}
\bibliography{resources}

\newpage
\input{appendix.tex}

\end{document}