### 2025-06-16 docs/paper/fig/architecture.tex
feat: add case studies

\begin{figure}
    \centering
    \begin{tikzpicture}[node distance=1.2cm and 1.8cm,
            input-block/.style={rectangle, draw, fill=red!10, text width=3.5cm, minimum height=1.2cm, align=center, rounded corners},
            output-block/.style={rectangle, draw, fill=green!10, text width=3.5cm, minimum height=1.2cm, align=center, rounded corners},
            block/.style={rectangle, draw, fill=blue!10, text width=3.5cm, minimum height=1.2cm, align=center, rounded corners},
            arrow/.style={thick,->,>=Stealth},
            tool/.style={rectangle, draw, fill=magenta!10, text width=3cm, minimum height=0.8cm, align=center, rounded corners},
            inv/.style={}
        ]

        % Input block
        \node[input-block] (input) {Config File};

        % Lucid blocks
        \node[block, right=of input] (estimator) {Estimator};
        \node[block, right=of estimator] (feature) {Feature map / Barrier template};
        \node[block, below=of feature] (optimizer) {Optimizer};
        \node[block, left=of optimizer] (verifier) {Verifier};
        \node[draw, dashed, inner sep=0.3cm, rounded corners,
        fit=(estimator)(feature)(optimizer)(verifier), label=above:{\textbf{Lucid}}] {};

        % Output block
        \node[output-block, left=of verifier] (output) {Barrier certificate};

        % Smt solvers
        \node[inv, below=2.5cm of verifier.north, anchor=north] (smtInv) {};
        \node[tool, left=0.3cm of smtInv] (dreal) {dReal};
        \node[tool, right=0.3cm of smtInv] (cvc5) {CVC5};
        \node[draw, dashed, inner sep=0.3cm, rounded corners,
        fit=(dreal)(cvc5), label=left:{\textbf{SMT Solver}}] {};

        % Arrows
        \draw[arrow] (input) -- (estimator);
        \draw[arrow] (estimator) -- (feature);
        \draw[arrow] (feature) -- (optimizer);
        \draw[arrow] (optimizer) -- (verifier);
        \draw[arrow] (verifier) -- (output);

        \draw[arrow] (cvc5.north) -- (verifier.south);
        \draw[arrow] (dreal.north) -- (verifier.south);

    \end{tikzpicture}
    \caption{General architecture of Lucid. \todo{Only dReal is supported at the moment}}
    \label{fig:architecture}
\end{figure}