### 2025-07-12 docs/paper/figures/steps.tex

\begin{figure}[ht]
    \centering
    \begin{tikzpicture}[node distance=0.6cm and 0.9cm,
            input-block/.style={rectangle, draw, fill=red!10, text width=3.5cm, minimum height=1.0cm, align=center, rounded corners},
            output-block/.style={rectangle, draw, fill=green!10, text width=3.5cm, minimum height=1.0cm, align=center, rounded corners},
            block/.style={rectangle, draw, fill=blue!10, text width=3.5cm, minimum height=1.0cm, align=center, rounded corners},
            arrow/.style={thick,->,>=Stealth},
            tool/.style={rectangle, draw, fill=green!20, text width=2cm, minimum height=0.8cm, align=center, rounded corners},
            inv/.style={}
        ]

        % Input block
        \node[input-block] (input) {Collect samples $\{(x_i, x^+_i)\}_{i=1}^N$};

        % Lucid blocks
        \node[block, right=of input] (estimator) {Fit the estimator $\hat{f}: \X \to \mathbb{R}^M$};
        \node[block, below=of estimator] (feature) {Construct the feature map $\phi: \X \to \mathbb{R}^M$};
        \node[block, below=of feature] (lattice) {Construct lattices on $\X, \X_0, \X_u$};
        \node[block, below=of lattice] (apply-lattice) {Apply the feature map to the lattices points};
        \node[block, below=of apply-lattice] (optimizer) {Optimize the LP};
        \node[draw, dashed, inner sep=0.3cm, rounded corners,
        fit=(estimator)(feature)(lattice)(apply-lattice)(optimizer), label=above:{\textbf{Lucid}}] {};

        % Output block
        \node[output-block, left=of optimizer] (output) {Barrier certificate};

        % Arrows
        \draw[arrow] (input) -- (estimator);
        \draw[arrow] (estimator) -- (feature);
        \draw[arrow] (feature) -- (lattice);
        \draw[arrow] (lattice) -- (apply-lattice);
        \draw[arrow] (apply-lattice) -- (optimizer);
        \draw[arrow] (optimizer) -- (output);

    \end{tikzpicture}
    \caption{Sequence of steps \lucid goes through to generate a barrier certificate.}
    \label{fig:steps}
\end{figure}