### 2025-07-30 docs/paper/appendix.tex
feat: update paper

% !TEX root =  main.tex
\appendix

% \section{Additional Theoretical Details}
% \paragraph{RKHS basics.}
% A symmetric function $k_\X:\X\times\X\rightarrow\R$ is called a (positive definite) \emph{kernel} (note the distinction from \emph{probability kernels}) if for all $N\in\N_{>0}$ we have $\sum_{i=1}^{N}\sum_{j=1}^{N}a_i a_j\allowbreak k_\X(x_i,x_j) \geq 0$ for $x_1,\ldots,x_N\in\X\subset{\R^n}$ and $a_1,\ldots,a_N\in\R$.
% A prominent example is the \emph{squared exponential} (SQExp) kernel \shortcite{Rasmussen2005GP,Kanagawa2018GPvsKernel}:
% \begin{equation}
%     k_\X(x,x') := \sigma_f^2 \exp\left( -\frac{1}{2} (x-x')\T \Sigma\, (x-x') \right),\quad \Sigma:=\mathrm{diag}(\sigma_l)^{-2},\label{eq:sqexp_kernel}
% \end{equation}
% with amplitude $\sigma_f^2\geq0$ and lengthscale coefficients $\sigma_l\in\R^n$.
% \new{For this work,} we assume that all kernels are bounded on their domain, i.e., $\E_{}[k_\X(x,x)]<\infty$, $x\in\X$.
% Given a kernel $k_\X$ on a non-empty set $\X$, there exists a unique corresponding
% \emph{reproducing kernel Hilbert space} (RKHS) $\Hilbert_{k_\X}$
% of functions $f:\X\rightarrow\R$ equipped with an inner product $\innerH{\cdotx}{\cdotx}{\Hilbert_{k_\X}}$
% with the celebrated \emph{reproducing property} such that for any function $f\in\Hilbert_{k_\X}$ and $x\in\X$ we have $f(x)=\innerH{f}{k_\X(\cdotx,x)}{\Hilbert_{k_\X}}$.
% Note that $k_\X(\cdotx,x):\X\rightarrow\Hilbert_{k_\X}$ is a real-valued function,
% which is also called an implicit \emph{canonical} \emph{embedding} or \emph{feature map} $\phi_\X$
% such that $k_\X(x,x')=\innerH{\phi_\X(x)}{\phi_\X(x')}{\Hilbert_{k_\X}}$ for all $x,x'\in\X$.
% For an RKHS $\Hilbert_{k_\X}$, we use the associated feature map $\phi_\X$ and kernel $k_\X$ interchangeably for ease of notation and comprehensibility.
% The inner product induces the norm $\norm{f}_{\Hilbert_{k_\X}}\!\!\!\!:=\!\!\sqrt{\smash[b]{\innerH{f}{f}{\Hilbert_{k_\X}}}}$ of the RKHS. 
% Throughout this paper, we assume that all RKHSs are \emph{separable}.
% Refer to the monograph by \citet{Berlinet2004RKHSProbStat} for a comprehensive study on RKHSs.
% Given $N$ i.i.d. samples $\hat X_N:=[\hat{x}_i]_{i=1}^N$ with $\hat{x}_i\in\X$, the
% \emph{Gram matrix} of $k_\X$ is given by
% $\new{K_{\hat{X}}^N}:=[k_\X(\hat{x}_i,\hat{x}_j)]_{i,j=1}^N.$
% Furthermore, we define the vector-valued function 
% $\new{k_{\hat{X}}^N}(x)  := [k_\X(x,\hat{x}_i)]_{i=1}^N.$

% \begin{definition}[Conditional Mean Embedding (CME)]\label{def:condMeanEmbed}
%     Given two RKHSs $\Hilbert_{k_\X}$ and $\Hilbert_{k_\Y}$ with the associated kernels $k_\X\colon\X\times\X\to\R$ and $k_\Y\colon\X\times\X\to\R$, the \emph{CME} of a probability kernel $\Tr\colon\X\times\borel{\Y}\rightarrow[0,1]$ is an $X$-measurable random variable taking values in $\Hilbert_{k_\Y}$, given by
%     \begin{equation*}
%         \cme_{k_\Y|k_\X}(\Tr)(\cdotx) := \E_{\Tr}[\phi_\Y(Y)\mid X=\cdotx].
%     \end{equation*}
% \end{definition}

\section{Linear Program}
The finitely-constrained LP discussed in Section~\ref{sec:ddbarriers} is presented. For this, the lattices $X_{{N}}\subset\X$, $\smash{\{ x_0^{1}, \ldots, x_0^{{N}_0} \}} \subset \X_0$, and $\smash{\{ x_u^{1}, \ldots, x_u^{{N}_u} \}} \subset \X_u$ of cardinality ${N}_0\in\N$ and ${N}_u\in\N$, respectively, are formed.
%
For given values of $\overline{\B}$, $\gamma$, and robustness radius $\mathcal{R}\geq0$, the following LP is obtained:
    \begin{equation*}
        \begin{alignedat}{3}
                  & \min_{\stackrel{b, c, \eta}{\Bmin_{{N}}^{\X_0}, \Bmax_{{N}}^{\X_u},\Bmax_{{N}}^\X,\Bmin_\Delta}}\hspace{-1.5em}  &                                            & \eta + cT,                                          &                       &
            %\label{eq:blackbox:linear_prog_objective}
            \\
                  & \text{subject to}
                  &                                                                                                                                & \phi_M(x_0^{i})\T b\leq\hat{\eta}, \quad &                                                     & i=1,\ldots,{N}_0,
            %\label{eq:blackbox:linear_prog_initial}
            \\
                  &                                                                                                                                &                                            & \phi_M(x_u^{i})\T b\geq\hat{\gamma},
            \quad &                                                                                                                                & i=1,\ldots,{N}_u,
            %\label{eq:blackbox:linear_prog_unsafe}
            \\
                  &                                                                                                                                &                                            & \phi_M(x^{i})\T(Hb - b) \leq \hat{\Delta},
            \quad &                                                                                                                                & i=1,\ldots,{N},
            %\label{eq:blackbox:linear_prog_kushner}
            \\
                  &                                                                                                                                &                                            & \phi_M(x^{i})\T b\geq \hat{\xi},
            \quad &                                                                                                                                & i=1,\ldots,{N},
            %\label{eq:blackbox:linear_prog_positive}
            \\
                  &                                                                                                                                &                                            & \Bmin_{{N}}^{\X_0}\leq\phi_M(x_0^{i})\T b,
            \quad &                                                                                                                                & i=1,\ldots,{N}_0,                                                                                                      \\
                  &                                                                                                                                &                                            & \Bmax_{{N}}^{\X_u}\geq\phi_M(x_u^{i})\T b,
            \quad &                                                                                                                                & i=1,\ldots,{N}_u,                                                                                                      \\
                  &                                                                                                                                &                                            & \Bmax_{{N}}^\X\geq\phi_M(x^{i})\T b,
            \quad &                                                                                                                                & i=1,\ldots,{N},                                                                                                        \\
                  &                                                                                                                                &                                            & \Bmin_\Delta\leq\phi_M(x^{i})\T (Hb - b),
            \quad &                                                                                                                                & i=1,\ldots,{N},                                                                                                        \\
                  &                                                                                                                                &                                            & c\geq 0,\,\gamma>\eta\geq 0,\, b\in\R^{2M+1},       &                       &
            \label{eq:blackbox:linear_prog}%
        \end{alignedat}
    \end{equation*}
with $\overline{\kappa}\geq\sigma_f$, $\overline{\B}\geq\norm{b}_2$, and constraint-tightening coefficients
    \begin{align*}
        \hat{\eta}   & := \frac{2}{C_{{N}}+1}\eta + \frac{C_{{N}}-1}{C_{{N}}+1}\Bmin_{{N}}^{\X_0},
                     \\                                                                                                                                    & \hat{\gamma} := \frac{2}{C_{{N}}+1}\gamma + \frac{C_{{N}}-1}{C_{{N}}+1}\Bmax_{{N}}^{\X_u}, \\
        \hat{\Delta} & := \frac{2}{C_{{N}}+1}\left(c - \mathcal{R}\overline{\B}\overline{\kappa}\right) + \frac{C_{{N}}-1}{C_{{N}}+1}\Bmin_\Delta,
                     \\                                                                                                                                     & \hat{\xi} := \frac{C_{{N}}-1}{C_{{N}}+1}\Bmax_{{N}}^\X.
    \end{align*}

    % \section{Proofs}
    % We have collected a series of proofs of the claims we made in the main text.

    % \section{Benchmarks}
    % \label{app:benchmarks}
    % Complete list of results for the benchmarks presented in \autoref{sec:benchmarks}.
    % The legend is as follows:
    % \textit{Output} indicates the result produced by the solver (i.e., optimal, unbounded, infeasible or unspecified),
    % \textit{Prec} indicates the number of bits used in the last floating-point number representation,
    % \textit{Ref} indicates the number of refinements,
    % and \textit{time} is the time in seconds.