### 2025-07-30 docs/paper/appendix.tex
feat: add appendix

% !TEX root =  main.tex
\appendix

% \section{Additional Theoretical Details}
% \paragraph{RKHS basics.}
% A symmetric function $k_\X:\X\times\X\rightarrow\R$ is called a (positive definite) \emph{kernel} (note the distinction from \emph{probability kernels}) if for all $N\in\N_{>0}$ we have $\sum_{i=1}^{N}\sum_{j=1}^{N}a_i a_j\allowbreak k_\X(x_i,x_j) \geq 0$ for $x_1,\ldots,x_N\in\X\subset{\R^n}$ and $a_1,\ldots,a_N\in\R$.
% A prominent example is the \emph{squared exponential} (SQExp) kernel \shortcite{Rasmussen2005GP,Kanagawa2018GPvsKernel}:
% \begin{equation}
%     k_\X(x,x') := \sigma_f^2 \exp\left( -\frac{1}{2} (x-x')\T \Sigma\, (x-x') \right),\quad \Sigma:=\mathrm{diag}(\sigma_l)^{-2},\label{eq:sqexp_kernel}
% \end{equation}
% with amplitude $\sigma_f^2\geq0$ and lengthscale coefficients $\sigma_l\in\R^n$.
% \new{For this work,} we assume that all kernels are bounded on their domain, i.e., $\E_{}[k_\X(x,x)]<\infty$, $x\in\X$.
% Given a kernel $k_\X$ on a non-empty set $\X$, there exists a unique corresponding
% \emph{reproducing kernel Hilbert space} (RKHS) $\Hilbert_{k_\X}$
% of functions $f:\X\rightarrow\R$ equipped with an inner product $\innerH{\cdotx}{\cdotx}{\Hilbert_{k_\X}}$
% with the celebrated \emph{reproducing property} such that for any function $f\in\Hilbert_{k_\X}$ and $x\in\X$ we have $f(x)=\innerH{f}{k_\X(\cdotx,x)}{\Hilbert_{k_\X}}$.
% Note that $k_\X(\cdotx,x):\X\rightarrow\Hilbert_{k_\X}$ is a real-valued function,
% which is also called an implicit \emph{canonical} \emph{embedding} or \emph{feature map} $\phi_\X$
% such that $k_\X(x,x')=\innerH{\phi_\X(x)}{\phi_\X(x')}{\Hilbert_{k_\X}}$ for all $x,x'\in\X$.
% For an RKHS $\Hilbert_{k_\X}$, we use the associated feature map $\phi_\X$ and kernel $k_\X$ interchangeably for ease of notation and comprehensibility.
% The inner product induces the norm $\norm{f}_{\Hilbert_{k_\X}}\!\!\!\!:=\!\!\sqrt{\smash[b]{\innerH{f}{f}{\Hilbert_{k_\X}}}}$ of the RKHS. 
% Throughout this paper, we assume that all RKHSs are \emph{separable}.
% Refer to the monograph by \citet{Berlinet2004RKHSProbStat} for a comprehensive study on RKHSs.
% Given $N$ i.i.d. samples $\hat X_N:=[\hat{x}_i]_{i=1}^N$ with $\hat{x}_i\in\X$, the
% \emph{Gram matrix} of $k_\X$ is given by
% $\new{K_{\hat{X}}^N}:=[k_\X(\hat{x}_i,\hat{x}_j)]_{i,j=1}^N.$
% Furthermore, we define the vector-valued function 
% $\new{k_{\hat{X}}^N}(x)  := [k_\X(x,\hat{x}_i)]_{i=1}^N.$

% \begin{definition}[Conditional Mean Embedding (CME)]\label{def:condMeanEmbed}
%     Given two RKHSs $\Hilbert_{k_\X}$ and $\Hilbert_{k_\Y}$ with the associated kernels $k_\X\colon\X\times\X\to\R$ and $k_\Y\colon\X\times\X\to\R$, the \emph{CME} of a probability kernel $\Tr\colon\X\times\borel{\Y}\rightarrow[0,1]$ is an $X$-measurable random variable taking values in $\Hilbert_{k_\Y}$, given by
%     \begin{equation*}
%         \cme_{k_\Y|k_\X}(\Tr)(\cdotx) := \E_{\Tr}[\phi_\Y(Y)\mid X=\cdotx].
%     \end{equation*}
% \end{definition}

\section{Linear Program}
The finitely-constrained LP discussed in Section~\ref{sec:ddbarriers} is presented. For this, the lattices $X_{{N}}\subset\X$, $\smash{\{ x_0^{1}, \ldots, x_0^{{N}_0} \}} \subset \X_0$, and $\smash{\{ x_u^{1}, \ldots, x_u^{{N}_u} \}} \subset \X_u$ of cardinality ${N}_0\in\N$ and ${N}_u\in\N$, respectively, are formed.
%
For given values of $\overline{\B}$, $\gamma$, and robustness radius $\mathcal{R}\geq0$, the following LP is obtained:
\begin{equation*}
      \begin{alignedat}{3}
                  & \min_{\stackrel{b, c, \eta}{\Bmin_{{N}}^{\X_0}, \Bmax_{{N}}^{\X_u},\Bmax_{{N}}^\X,\Bmin_\Delta}}\hspace{-1.5em} &                                          & \eta + cT,                                    &                   &
            %\label{eq:blackbox:linear_prog_objective}
            \\
                  & \text{subject to}
                  &                                                                                                                 & \phi_M(x_0^{i})\T b\leq\hat{\eta}, \quad &                                               & i=1,\ldots,{N}_0,
            %\label{eq:blackbox:linear_prog_initial}
            \\
                  &                                                                                                                 &                                          & \phi_M(x_u^{i})\T b\geq\hat{\gamma},
            \quad &                                                                                                                 & i=1,\ldots,{N}_u,
            %\label{eq:blackbox:linear_prog_unsafe}
            \\
                  &                                                                                                                 &                                          & \phi_M(x^{i})\T(Hb - b) \leq \hat{\Delta},
            \quad &                                                                                                                 & i=1,\ldots,{N},
            %\label{eq:blackbox:linear_prog_kushner}
            \\
                  &                                                                                                                 &                                          & \phi_M(x^{i})\T b\geq \hat{\xi},
            \quad &                                                                                                                 & i=1,\ldots,{N},
            %\label{eq:blackbox:linear_prog_positive}
            \\
                  &                                                                                                                 &                                          & \Bmin_{{N}}^{\X_0}\leq\phi_M(x_0^{i})\T b,
            \quad &                                                                                                                 & i=1,\ldots,{N}_0,                                                                                              \\
                  &                                                                                                                 &                                          & \Bmax_{{N}}^{\X_u}\geq\phi_M(x_u^{i})\T b,
            \quad &                                                                                                                 & i=1,\ldots,{N}_u,                                                                                              \\
                  &                                                                                                                 &                                          & \Bmax_{{N}}^\X\geq\phi_M(x^{i})\T b,
            \quad &                                                                                                                 & i=1,\ldots,{N},                                                                                                \\
                  &                                                                                                                 &                                          & \Bmin_\Delta\leq\phi_M(x^{i})\T (Hb - b),
            \quad &                                                                                                                 & i=1,\ldots,{N},                                                                                                \\
                  &                                                                                                                 &                                          & c\geq 0,\,\gamma>\eta\geq 0,\, b\in\R^{2M+1}, &                   &
            \label{eq:blackbox:linear_prog}%
      \end{alignedat}
\end{equation*}
with $\overline{\kappa}\geq\sigma_f$, $\overline{\B}\geq\norm{b}_2$, and constraint-tightening coefficients
\begin{align*}
      \hat{\eta}   & := \frac{2}{C_{{N}}+1}\eta + \frac{C_{{N}}-1}{C_{{N}}+1}\Bmin_{{N}}^{\X_0},
      \\                                                                                                                                    & \hat{\gamma} := \frac{2}{C_{{N}}+1}\gamma + \frac{C_{{N}}-1}{C_{{N}}+1}\Bmax_{{N}}^{\X_u}, \\
      \hat{\Delta} & := \frac{2}{C_{{N}}+1}\left(c - \mathcal{R}\overline{\B}\overline{\kappa}\right) + \frac{C_{{N}}-1}{C_{{N}}+1}\Bmin_\Delta,
      \\                                                                                                                                     & \hat{\xi} := \frac{C_{{N}}-1}{C_{{N}}+1}\Bmax_{{N}}^\X.
\end{align*}

% \section{Proofs}
% We have collected a series of proofs of the claims we made in the main text.

\section{Installation}
Provided \texttt{Python>=3.8} is already present, \pylucid can be installed with the command
\begin{lstlisting}[language=bash,numbers=none,xleftmargin=0em]
pip install pylucid[gui,plot] --index-url \
  https://gitlab.com/api/v4/projects/71977529/packages/pypi/simple
\end{lstlisting}
There are a few limitation some of the dependencies introduce, which may preclude the ability of using certain features on not supported platforms.
Formal verification of the barrier requires the \dreal SMT solver, which can only be installed on Linux and non-ARM macOS.
Support for \highs \gls{lp} is not available on Windows at the moment.
The \gurobi solver must be installed beforehand, and requires a license for commercial use, but is free for academic purposes.

For more details on the installation process, please refer to the online documentation at \url{https://lucidtoolsource.gitlab.io/lucid/md_docs_2Pylucid.html}.

\section{Benchmarks}
\label{app:benchmarks}
Complete list of results for the benchmarks presented in Section~\ref{sec:experiments}.
The legend is as follows:
\textit{Output} indicates the result produced by the solver (i.e., optimal, unbounded, infeasible or unspecified),
\textit{Prec} indicates the number of bits used in the last floating-point number representation,
\textit{Ref} indicates the number of refinements,
and \textit{time} is the time in seconds.

All benchmarks were run on a Windows 10 machine with an AMD Ryzen 9 5950X 16-Core Processor @ 3.40 GHz, and 64 GB of RAM, and all runs had the random seed set to $42$ to ensure reproducibility.
The scripts responsible for running the benchmarks and \hp tuning are available in the \texttt{benchmarks/integration} folder of the repository.
Note that a version of \texttt{Python>=3.9} is required to run the scripts, as some dependencies used to track the benchmarks' metrics across multiple runs are not compatible with \texttt{Python 3.8}.
The \hp tuning of the \estimator was performed with the \texttt{LbfgsTuner}, bounding the value of $\sigma_l$ between $[10^{-5}, 10^5]$, and refined with the \texttt{GridSearchTuner}.
The implementation can be found in the \texttt{benchmarks/integration/hp\_tuning.py} script. 
Note that we used the \gurobi optimiser.
Other optimizers, such as \alglib or \highs, can be used instead, but may yield different results, especially in terms of performance.

\subsection{Linear}

We consider the following system:
\begin{equation*}
    \begin{bmatrix}
        {x}_{t+1}
    \end{bmatrix}
    = \begin{bmatrix}
        0.5 {x}_{t}
    \end{bmatrix} + w_t,
\end{equation*}
where $w_t\sim\mathcal{N}(\cdotx\vert 0,0.01I_1)$.
Given
\begin{align*}
    &\X = [-1, 1] \\
    &\X_0 = [-0.5, 0.5] \\
    &\X_U = [-1, -0.9] \cup [0.9, 1],
\end{align*}
we want to ensure that the system, starting in $\X_0$, does not enter the unsafe regions $\X_U$ within $T=15$ time steps.
The complete configuration for the linear example benchmark is shown in Listing~\ref{lst:linear}.
\lstinputlisting[language=yaml,caption={Configuration for linear example},captionpos=b,label={lst:linear}]{code/linear.yaml}

\subsection{Barrier 2}

We consider the following system:
\begin{equation*}
    \begin{bmatrix}
        {x}_{1, t+1} \\
        {x}_{2, t+1}
    \end{bmatrix}
    = \begin{bmatrix}
        {x}_{2, t} - 1 + e^{-x_{1, t}} \\
        -\sin^2(x_{1, t})
    \end{bmatrix} + w_t,
\end{equation*}
where $w_t\sim\mathcal{N}(\cdotx\vert 0,0.01I_2)$.
Given
\begin{align*}
    &\X = [ -2, 2 ] \times [ -2 , 2 ] \\
    &\X_0 = \{ [x_1, x_2] : (x_1 + 0.5)^2 + (x_2 - 0.5)^2 \leq 0.4 \} \\
    &\X_U = \{ [x_1, x_2] : (x_1 - 0.7)^2 + (x_2 + 0.7)^2 \leq 0.3 \},
\end{align*}
we want to ensure that the system, starting in $\X_0$, does not enter the unsafe regions $\X_U$ within $T=5$ time steps.
The complete configuration for the \barrII benchmark is shown in Listing~\ref{lst:barrier2}.
\lstinputlisting[language=yaml,caption={Configuration for \barrII},captionpos=b,label={lst:barrier2}]{code/barrier2.yaml}

\subsection{Barrier 3}

We consider the following system:
\begin{equation*}
    \begin{bmatrix}
        {x}_{1, t+1} \\
        {x}_{2, t+1}
    \end{bmatrix}
    = \begin{bmatrix}
        {x}_{2, t} \\
        \frac{1}{3} {x}^3_{1, t} - {x}_{1,t} - {x}_{2,t}
    \end{bmatrix} + w_t,
\end{equation*}
where $w_t\sim\mathcal{N}(\cdotx\vert 0,0.01I_2)$.
Given
\begin{align*}
    &\X = [ -3, 2.5 ] \times [ -2 , 1 ] \\
    &\X_0 = [ 1 , 2 ] \times [ -0.7 , 0.3 ] \cup [ -1.8 , -1.4 ] \times [ -0.1 , 0.1 ] \\
    &\qquad \cup [-1.4, -1.2] \times [-0.5 , 0.1] \\
    &\X_U = [ 0.4 , 0.6 ] \times [ 0.2 , 0.6 ] \cup [ 0.6 , 0.7 ] \times [ 0.2 , 0.4 ],
\end{align*}
we want to ensure that the system, starting in $\X_0$, does not enter the unsafe regions $\X_U$ within $T=5$ time steps.
The complete configuration for the \barrIII benchmark is shown in Listing~\ref{lst:barrier3}.
\lstinputlisting[language=yaml,caption={Configuration for \barrIII},captionpos=b,label={lst:barrier3}]{code/barrier3.yaml}

\subsection{Overtaking}

We consider a scenario where an autonomous vehicle (AV) controlled by a \gls{nn} is overtaking another vehicle.
The dynamics of the ego vehicle are given by Dubin's car model \EC{does it need a cit or is it just textbook} by adding the noise vector $w = \begin{bmatrix}w_t^1 & w_t^2 & w_t^3\end{bmatrix}$ where each component is drawn from a zero-mean Gaussian with standard deviation $0.01$, $0.01$, and $0.001$, respectively.
The steering wheel angle is supplied by the \gls{nn} controller and we travel at a fixed velocity $v=1$.
Given
\todo{correct the sets}
\begin{align*}
    &\X = [ -3, 2.5 ] \times [ -2 , 1 ] \\
    &\X_0 = [ 1 , 2 ] \times [ -0.7 , 0.3 ] \cup [ -1.8 , -1.4 ] \times [ -0.1 , 0.1 ] \\
    &\qquad \cup [-1.4, -1.2] \times [-0.5 , 0.1] \\
    &\X_U = [ 0.4 , 0.6 ] \times [ 0.2 , 0.6 ] \cup [ 0.6 , 0.7 ] \times [ 0.2 , 0.4 ],
\end{align*}
we want to ensure that the system, starting in $\X_0$, does not enter the unsafe regions $\X_U$ within $T=5$ time steps.

The complete configuration for the \overtaking benchmark is shown in Listing~\ref{lst:overtaking}.
\lstinputlisting[language=yaml,caption={Configuration for \overtaking},captionpos=b,label={lst:overtaking}]{code/overtaking.yaml}