### 2025-08-11 docs/paper/appendix.tex
feat: submitted

% !TEX root =  main.tex
\appendix
\onecolumn

% \section{Additional Theoretical Details}
% \paragraph{RKHS basics.}
% A symmetric function $k_\X:\X\times\X\rightarrow\R$ is called a (positive definite) \emph{kernel} (note the distinction from \emph{probability kernels}) if for all $N\in\N_{>0}$ we have $\sum_{i=1}^{N}\sum_{j=1}^{N}a_i a_j\allowbreak k_\X(x_i,x_j) \geq 0$ for $x_1,\ldots,x_N\in\X\subset{\R^n}$ and $a_1,\ldots,a_N\in\R$.
% A prominent example is the \emph{squared exponential} (SQExp) kernel \shortcite{Rasmussen2005GP,Kanagawa2018GPvsKernel}:
% \begin{equation}
%     k_\X(x,x') := \sigma_f^2 \exp\left( -\frac{1}{2} (x-x')\T \Sigma\, (x-x') \right),\quad \Sigma:=\mathrm{diag}(\sigma_l)^{-2},\label{eq:sqexp_kernel}
% \end{equation}
% with amplitude $\sigma_f^2\geq0$ and lengthscale coefficients $\sigma_l\in\R^n$.
% \new{For this work,} we assume that all kernels are bounded on their domain, i.e., $\E_{}[k_\X(x,x)]<\infty$, $x\in\X$.
% Given a kernel $k_\X$ on a non-empty set $\X$, there exists a unique corresponding
% \emph{reproducing kernel Hilbert space} (RKHS) $\Hilbert_{k_\X}$
% of functions $f:\X\rightarrow\R$ equipped with an inner product $\innerH{\cdotx}{\cdotx}{\Hilbert_{k_\X}}$
% with the celebrated \emph{reproducing property} such that for any function $f\in\Hilbert_{k_\X}$ and $x\in\X$ we have $f(x)=\innerH{f}{k_\X(\cdotx,x)}{\Hilbert_{k_\X}}$.
% Note that $k_\X(\cdotx,x):\X\rightarrow\Hilbert_{k_\X}$ is a real-valued function,
% which is also called an implicit \emph{canonical} \emph{embedding} or \emph{feature map} $\phi_\X$
% such that $k_\X(x,x')=\innerH{\phi_\X(x)}{\phi_\X(x')}{\Hilbert_{k_\X}}$ for all $x,x'\in\X$.
% For an RKHS $\Hilbert_{k_\X}$, we use the associated feature map $\phi_\X$ and kernel $k_\X$ interchangeably for ease of notation and comprehensibility.
% The inner product induces the norm $\norm{f}_{\Hilbert_{k_\X}}\!\!\!\!:=\!\!\sqrt{\smash[b]{\innerH{f}{f}{\Hilbert_{k_\X}}}}$ of the RKHS. 
% Throughout this paper, we assume that all RKHSs are \emph{separable}.
% Refer to the monograph by \citet{Berlinet2004RKHSProbStat} for a comprehensive study on RKHSs.
% Given $N$ i.i.d. samples $\hat X_N:=[\hat{x}_i]_{i=1}^N$ with $\hat{x}_i\in\X$, the
% \emph{Gram matrix} of $k_\X$ is given by
% $\new{K_{\hat{X}}^N}:=[k_\X(\hat{x}_i,\hat{x}_j)]_{i,j=1}^N.$
% Furthermore, we define the vector-valued function 
% $\new{k_{\hat{X}}^N}(x)  := [k_\X(x,\hat{x}_i)]_{i=1}^N.$

% \begin{definition}[Conditional Mean Embedding (CME)]\label{def:condMeanEmbed}
%     Given two RKHSs $\Hilbert_{k_\X}$ and $\Hilbert_{k_\Y}$ with the associated kernels $k_\X\colon\X\times\X\to\R$ and $k_\Y\colon\X\times\X\to\R$, the \emph{CME} of a probability kernel $\Tr\colon\X\times\borel{\Y}\rightarrow[0,1]$ is an $X$-measurable random variable taking values in $\Hilbert_{k_\Y}$, given by
%     \begin{equation*}
%         \cme_{k_\Y|k_\X}(\Tr)(\cdotx) := \E_{\Tr}[\phi_\Y(Y)\mid X=\cdotx].
%     \end{equation*}
% \end{definition}

\section{Linear Program}
The finitely-constrained LP discussed in Section~\ref{sec:ddbarriers} is presented. For this, the lattices $\smash{\{ x^{1}, \ldots, x^{{N}} \}}\subset\X$, $\smash{\{ x_0^{1}, \ldots, x_0^{{N}_0} \}} \subset \X_0$, and $\smash{\{ x_U^{1}, \ldots, x_U^{N_U} \}} \subset \X_U$ of cardinality $N\in\N$, ${N}_0\in\N$, and $N_U\in\N$, respectively, are formed.
%
For given values of $\overline{\B}$, $\gamma$, and robustness radius $\mathcal{R}\geq0$, the following LP is obtained:
\begin{equation*}
      \begin{alignedat}{3}
            & \min_{\stackrel{b, c, \eta}{\Bmin_{{N}}^{\X_0}, \Bmax_{{N}}^{\X_U},\Bmax_{{N}}^\X,\Bmin_\Delta}} &                                          & \eta + cT,                                    &                   &
            %\label{eq:blackbox:linear_prog_objective}
            \\
            & \text{subject to}
            &                                                                                                  & \phi_M(x_0^{i})\T b\leq\hat{\eta}, \quad &                                               & i=1,\ldots,{N}_0,
            %\label{eq:blackbox:linear_prog_initial}
            \\
            &                                                                                                  &                                          & \phi_M(x_U^{i})\T b\geq\hat{\gamma},
            \quad &                                                                                                  & i=1,\ldots,N_U,
            %\label{eq:blackbox:linear_prog_unsafe}
            \\
            &                                                                                                  &                                          & \phi_M(x^{i})\T(Hb - b) \leq \hat{\Delta},
            \quad &                                                                                                  & i=1,\ldots,{N},
            %\label{eq:blackbox:linear_prog_kushner}
            \\
            &                                                                                                  &                                          & \phi_M(x^{i})\T b\geq \hat{\xi},
            \quad &                                                                                                  & i=1,\ldots,{N},
            %\label{eq:blackbox:linear_prog_positive}
            \\
            &                                                                                                  &                                          & \Bmin_{{N}}^{\X_0}\leq\phi_M(x_0^{i})\T b,
            \quad &                                                                                                  & i=1,\ldots,{N}_0,                                                                                              \\
            &                                                                                                  &                                          & \Bmax_{{N}}^{\X_U}\geq\phi_M(x_U^{i})\T b,
            \quad &                                                                                                  & i=1,\ldots,N_U,                                                                                              \\
            &                                                                                                  &                                          & \Bmax_{{N}}^\X\geq\phi_M(x^{i})\T b,
            \quad &                                                                                                  & i=1,\ldots,{N},                                                                                                \\
            &                                                                                                  &                                          & \Bmin_\Delta\leq\phi_M(x^{i})\T (Hb - b),
            \quad &                                                                                                  & i=1,\ldots,{N},                                                                                                \\
            &                                                                                                  &                                          & c\geq 0,\,\gamma>\eta\geq 0,\, b\in\R^{2M+1}, &                   &
            \label{eq:blackbox:linear_prog}%
      \end{alignedat}
\end{equation*}
with $\overline{\kappa}\geq\sigma_f$, $\overline{\B}\geq\norm{b}_2$, and constraint-tightening coefficients
\begin{align*}
      \hat{\eta}   & := \frac{2}{C_{{N}}+1}\eta + \frac{C_{{N}}-1}{C_{{N}}+1}\Bmin_{{N}}^{\X_0},
                   &                                                                                                                             & \hat{\gamma} := \frac{2}{C_{{N}}+1}\gamma + \frac{C_{{N}}-1}{C_{{N}}+1}\Bmax_{{N}}^{\X_U}, \\
      \hat{\Delta} & := \frac{2}{C_{{N}}+1}\left(c - \mathcal{R}\overline{\B}\overline{\kappa}\right) + \frac{C_{{N}}-1}{C_{{N}}+1}\Bmin_\Delta,
                   &                                                                                                                             & \hat{\xi} := \frac{C_{{N}}-1}{C_{{N}}+1}\Bmax_{{N}}^\X.
\end{align*}

% \section{Proofs}
% We have collected a series of proofs of the claims we made in the main text.

\section{Tool Technical Details}

\subsection{Installation}
The recommended way to install \pylucid is via the \texttt{pip} package manager, using one of the pre-built wheels,
hence avoiding the time-consuming need to compile the code from source.
Provided \texttt{Python>=3.8} is already present on the system, \pylucid is installed using a single command:
%
\begin{lstlisting}[language=bash,numbers=none,xleftmargin=0em,backgroundcolor=\color{ipython_bg}]
pip install "pylucid[gui,plot]" --index-url https://gitlab.com/api/v4/projects/71977529/packages/pypi/simple
\end{lstlisting}
%
Note that installing \pylucid this way requires \gurobi to be installed beforehand, which can be obtained for free from the official website.
A valid license is not required unless the \gurobi solver is used.
If no wheel is available for the desired platform, \pylucid can be installed from source by running the following commands:
%
\begin{lstlisting}[language=bash,numbers=none,xleftmargin=0em,backgroundcolor=\color{ipython_bg}]
git clone https://gitlab.com/lucidtoolsource/lucid.git
cd lucid
pip install ".[gui,plot]"
\end{lstlisting}
%
Note that this requires \texttt{Python>=3.8} and \texttt{Bazel>=8.0} to be installed on the system.
For more details on the installation process, please refer to the online documentation at \url{https://lucidtoolsource.gitlab.io/lucid/md_docs_2Pylucid.html}.

\subsection{Containerized Installation}

We also provide a pre-built Docker image that can be used to run \pylucid in a containerized environment, making Docker the only dependency.
To run the image with the desired configuration, the following command can be used:
\begin{lstlisting}[language=bash,numbers=none,xleftmargin=0em,backgroundcolor=\color{ipython_bg}]
docker run --name lucid -it --rm \
  -v/path/to/conf.py:/config \
  registry.gitlab.com/lucidtoolsource/lucid:latest /config/conf.py
\end{lstlisting}
Alternatively, the container can also run the GUI by changing the command to
\begin{lstlisting}[language=bash,numbers=none,xleftmargin=0em,backgroundcolor=\color{ipython_bg}]
docker run --name lucid -it --rm \
  -p 3661:3661 --entrypoint pylucid-gui \
  registry.gitlab.com/lucidtoolsource/lucid:latest
\end{lstlisting}
The GUI will be accessible at \url{http://localhost:3661}.
In both cases, to use the \gurobi solver, a \gurobi Web License Service (WLS) license\footnote{See \url{https://www.gurobi.com/features/web-license-service/}} will have to be mounted in the container.

\subsection{Platform Support}
Some of \pylucid's dependencies are not supported on all platforms, which may preclude the ability of using certain features.
For instance, optional formal verification of the barrier requires the \dreal SMT solver, which can be installed only on Linux and non-ARM macOS.
The \highs \gls{lp} solver cannot be used on Windows at the moment.
The \gurobi solver must be installed separately beforehand and requires a licence, which is free for academic use.
Using the Docker container sidesteps most of these issues, with the only requirement being a CPU with x86-64 architecture.

\subsection{Configuration Formats}
\pylucid provides extensive support for various ways of specifying configurations, which can be defined as a \texttt{.yaml} file (recommended), \texttt{.json} file,
generated via a \texttt{.py} file returning a \texttt{Configuration} object, or provided directly via command line arguments.
Moreover, system transition data can be either embedded directly in the configuration file or provided separately in the form of \texttt{.csv}, \texttt{.mat}, \texttt{.npy}, or \texttt{.npz} files.
For example, the following configurations are equivalent:

\begin{minipage}{.45\columnwidth}
      \lstinputlisting[language=yaml,style={bgnonumbers},caption={YAML configuration.},captionpos=b,label={lst:configuration.yaml},backgroundcolor=\color{ipython_bg}]{code/configuration.yaml}
\end{minipage}\hfill
\begin{minipage}{.45\columnwidth}
      \lstinputlisting[language=json,style={bgnonumbers},caption={JSON configuration.},captionpos=b,label={lst:configuration.json}]{code/configuration.json}
\end{minipage}
\begin{minipage}{.45\columnwidth}
      \lstinputlisting[language=iPython,style={bgnonumbers},caption={Python configuration.},captionpos=b,label={lst:configuration.py}]{code/configuration.py}
\end{minipage}\hfill
\begin{minipage}{.45\columnwidth}
      \lstinputlisting[language=bash,style={bgnonumbers},caption={Command line configuration assuming that the transition samples have been stored as \texttt{.csv} files.},captionpos=b,label={lst:configuration.sh}]{code/configuration.sh}
\end{minipage}

\subsection{Graphical User Interface}

\pylucid provides a \gls{gui} to guide the user in the creation of the scenario configuration and presenting the results in an intuitive format.
Running \lstinline|pylucid-gui| will open a browser tab with the interface shown in
%Figures~\ref{fig:gui-left}--\ref{fig:gui-right}, 
Figure~\ref{fig:gui-full},
while a local server will listen for requests coming from the \gls{gui}, computing and returning the results.
Following the same enumeration as the arrows in the figure, we provide a list of the main components of the \gls{gui}:

\begin{figure}
      \centering
      \shadowimage[width=16.5cm]{figures/frontend-highlight.png}
      \caption{
            \pylucid's \gls{gui} as presented in the browser.
            The left half of the page allows the user to provide the transition samples and the safety specification to verify.
            The right half of the page shows the results produced by \lucid, including a plot of the computed barrier function and detailed logs.}
      \label{fig:gui-full}
\end{figure}

\begin{enumerate}
      \item It is easy to import or export configurations in JSON format.
            In the \textit{Import} menu, the user can paste the configuration or load it from a file.
            The experiments discussed in this paper are also available and can be selected from a dropdown list.
            The \textit{Export} menu can be used to copy the current configuration to the clipboard or download it locally.
      \item The dimension $n$ of the system is specified. Here, $n=1$.
      \item  The user can provide transition samples directly by copy--pasting them into the text fields in the Data tab.
            Alternatively, if a model describing the system’s expected behavior is available, the user can specify it in the Model tab.
            Lucid will then automatically generate a set of transitions from the model to be used in the computation of the barrier function.
            Note that this model-based generation is mainly intended for reference and convenience.
      \item The $x^1,\ldots,x^N$ samples are specified in CSV format. They can be edited manually or loaded from a file.
      \item The $x^1_+,\ldots,x^N_+$ samples are specified analogously.
      \item The state space $\X$ is defined by a \texttt{RectSet}. Here, $\X=[-1,1]$.
      \item The initial set $\X_0$ is defined from one of the available type options. A \texttt{MultiSet} can be defined via the \textit{+} button. Here, $\X_0=[-0.5,0.5]$.
      \item The unsafe set $\X_U$ is defined analogously. Here, $\X_U=[-1, -0.9] \cup [0.9, 1]$.
      \item The \textit{Preview} button visualizes the expected stochastic behavior of the black-box system, predicted from the transition samples or alternatively as specified in the \textit{Model} tab,
            alongside the bounding boxes of the state space $\X$, the initial set $\X_0$, and the unsafe set $\X_U$.
      \item The \textit{Submit} button starts the computation of the barrier function with the current configuration.
      \item The plot of the transition function or the computed barrier function.
      \item The numerical results of the computation, including the lower bound on the safety probability, values of $\eta$ and $c$, and the computation time.
            If a model of the expected system behavior is provided, it is also possible to formally verify the computed barrier function via \dreal.
      \item The logs provide information on the progress of the computation, updated in real-time.
            The verbosity of the logs can be adjusted in the \textit{Advanced} options.
      \item The main \gls{gui} displays the most important configuration options.
            More advanced options can be accessed via the \textit{Advanced} button.
            This includes the kernel \hp, time horizon, number of frequencies and lattice size, \gls{lp} solver selection, etc.
\end{enumerate}

\section{Benchmarks}
\label{app:benchmarks}
We provide a more detailed description of the benchmarks presented in Section~\ref{sec:experiments}, including the corresponding configuration files and an extended list of experiments.
All benchmarks were run on a Windows 10 machine with an AMD Ryzen 9 5950X 16-Core Processor @ 3.40 GHz, NVIDIA GeForce RTX 3090 GPU, and 64 GB of RAM.
To ensure reproducibility, all runs had the random seed set to $42$.
We ascertained that the variance of the safety probability is low across different seeds, and can be reduced further by increasing the number of samples used to fit the \estimator.
The scripts for running the benchmarks and tuning the \hp are available in the \texttt{benchmarks/integration/} folder of the repository.
Note that \texttt{Python>=3.9} is required to run the scripts, as some dependencies used to track the benchmarks' metrics across multiple runs are not compatible with Python 3.8.
The hyperparameter tuning of the \estimator was performed with the \texttt{LbfgsTuner}, bounding the value of $\sigma_l$ between $[10^{-5}, 10^5]$, and refined with the \texttt{GridSearchTuner}.
We use $N=1000$ sample transitions and the \gurobi solver for all experiments;
Other optimizers, such as \alglib or \highs, can be used instead, but may yield different results, especially in terms of performance.

\subsection{Linear}

We consider a black-box system with the following dynamics:
\begin{equation*}
      {x}_{t+1} = 0.5 {x}_{t} + w_t,
\end{equation*}
where $w_t\sim\mathcal{N}(\cdotx\vert 0,0.01I_1)$.
The data sampled from this black-box system is used as input for \lucid.
Given $\X = [-1, 1]$, $\X_0 = [-0.5, 0.5]$, and $\X_U = [-1, -0.9] \cup [0.9, 1]$,
we want to ensure that the system, starting in $\X_0$, does not enter the unsafe regions $\X_U$ within $T=15$ time steps.
We set the kernel \hp to $\sigma_f=15$, $\sigma_l=1.2$, and $\lambda=0.00001$.
The complete configuration for the linear example benchmark is shown in Listing~\ref{lst:linear}, with a satisfying barrier plotted in Figure~\ref{fig:example}.
\lstinputlisting[language=yaml,style={bgnonumbers},caption={Configuration for the linear example.},captionpos=b,label={lst:linear},backgroundcolor=\color{ipython_bg}]{code/linear.yaml}

\subsection{Barrier 2}

We consider a black-box system with the following nonlinear dynamics:
\begin{equation*}
      \begin{bmatrix}
            {x}_{1, t+1} \\
            {x}_{2, t+1}
      \end{bmatrix}
      = \begin{bmatrix}
            {x}_{1, t} \\
            {x}_{2, t}
      \end{bmatrix} + 0.1 \begin{bmatrix}
            {x}_{2, t} - 1 + e^{-x_{1, t}} \\
            -\sin^2(x_{1, t})
      \end{bmatrix} + w_t,
\end{equation*}
where $w_t\sim\mathcal{N}(\cdotx\vert 0,0.01I_2)$.
The data sampled from this black-box system is used as input for \lucid.
Given
\begin{align*}
       & \X = [ -2, 2 ] \times [ -2 , 2 ]   ,                              \\
       & \X_0 = \{ [x_1, x_2] : (x_1 + 0.5)^2 + (x_2 - 0.5)^2 \leq 0.4 \}, \\
       & \X_U = \{ [x_1, x_2] : (x_1 - 0.7)^2 + (x_2 + 0.7)^2 \leq 0.3 \},
\end{align*}
we want to ensure that the system, starting in $\X_0$, does not enter the unsafe regions $\X_U$ within $T=5$ time steps.
The system dynamics and safety specification are visualized in Figure~\ref{fig:model-barrier2}.
The kernel \hp are set to $\sigma_f=1$, $\sigma_l=[2.50, 3.78]$, and $\lambda=\smash{10^{-6}}$.
The complete configuration for the \barrII benchmark is shown in Listing~\ref{lst:barrier2}, with a satisfying barrier is plotted in Figure~\ref{fig:CSBarr2}.
A list of experiments using different combinations of frequencies and lattice sizes, showing their impact on performance and final result, is presented in Table~\ref{tab:results-barrier2}.
\lstinputlisting[language=yaml,style={bgnonumbers},caption={Configuration for \barrII.},captionpos=b,label={lst:barrier2}]{code/barrier2.yaml}

\begin{figure}[ht]
      \centering
      \input{figures/model-barrier2-noise.pgf}
      \caption{Visualization of the stochastic system behavior (based on 10 random transitions from an $100\times 100$ grid of initial states) and safety specification for the \barrII benchmark.}
      \label{fig:model-barrier2}
\end{figure}

\begin{table}[tb]
      \centering
      \begin{tabular}{ccccccc}
            \toprule
            \textbf{Freq.} & \textbf{Lattice} & $\eta$ & $\gamma$ & $c$  & \textbf{Runtime} & \textbf{Safety} \\
                           & \textbf{Size}    &        &          &      & [mm:ss]          & \textbf{Prob.}  \\ % Units
            \midrule
            17             & $700^2$          & 0.11   & 2        & 0.12 & 85:50            & 63.46\%         \\
            16             & $700^2$          & 0.11   & 2        & 0.12 & 57:15            & 63.24\%         \\
            18             & $600^2$          & 0.13   & 2        & 0.12 & 59:23            & 62.17\%         \\
            17             & $600^2$          & 0.13   & 2        & 0.12 & 43:53            & 61.93\%         \\
            16             & $600^2$          & 0.12   & 2        & 0.12 & 33:25            & 61.78\%         \\
            10             & $800^2$          & 0.11   & 2        & 0.13 & 15:27            & 60.57\%         \\
            18             & $500^2$          & 0.14   & 2        & 0.13 & 33:56            & 60.12\%         \\
            17             & $500^2$          & 0.14   & 2        & 0.13 & 24:10            & 59.95\%         \\
            16             & $500^2$          & 0.14   & 2        & 0.13 & 20:44            & 59.90\%         \\
            10             & $700^2$          & 0.11   & 2        & 0.13 & 10:47            & 59.80\%         \\
            9              & $800^2$          & 0.11   & 2        & 0.13 & 10:39            & 59.59\%         \\
            9              & $700^2$          & 0.12   & 2        & 0.13 & 7:14             & 58.89\%         \\
            10             & $600^2$          & 0.12   & 2        & 0.13 & 7:20             & 58.80\%         \\
            9              & $600^2$          & 0.12   & 2        & 0.14 & 5:22             & 57.98\%         \\
            10             & $500^2$          & 0.13   & 2        & 0.14 & 5:05             & 57.53\%         \\
            18             & $400^2$          & 0.17   & 2        & 0.13 & 19:09            & 57.16\%         \\
            15             & $400^2$          & 0.15   & 2        & 0.14 & 10:01            & 56.87\%         \\
            9              & $500^2$          & 0.13   & 2        & 0.14 & 3:40             & 56.82\%         \\
            10             & $400^2$          & 0.15   & 2        & 0.14 & 3:22             & 55.78\%         \\
            9              & $400^2$          & 0.15   & 2        & 0.14 & 2:20             & 55.25\%         \\
            11             & $300^2$          & 0.18   & 2        & 0.15 & 2:25             & 53.25\%         \\
            13             & $300^2$          & 0.18   & 2        & 0.15 & 3:53             & 53.16\%         \\
            10             & $300^2$          & 0.17   & 2        & 0.15 & 1:54             & 53.13\%         \\
            9              & $300^2$          & 0.17   & 2        & 0.15 & 1:25             & 52.86\%         \\
            18             & $300^2$          & 0.20   & 2        & 0.14 & 12:04            & 52.58\%         \\
            20             & $300^2$          & 0.22   & 2        & 0.14 & 16:48            & 51.93\%         \\
            \bottomrule
      \end{tabular}
      \caption{
            \barrII results, sorted by the last column. For each combination of number of frequencies, $M$, and lattice size (i.e., number of lattice points per dimension), we report the values of $c$, $\gamma$, $\lambda$, the runtime, and the achieved lower bound on the safety probability.}
      \label{tab:results-barrier2}
\end{table}

\subsection{Barrier 3}

We consider a black-box system that has the following dynamics:
\begin{equation*}
      \begin{bmatrix}
            {x}_{1, t+1} \\
            {x}_{2, t+1}
      \end{bmatrix}
      = \begin{bmatrix}
            {x}_{1, t} \\
            {x}_{2, t}
      \end{bmatrix} + 0.1 \begin{bmatrix}
            {x}_{2, t} \\
            \frac{1}{3} {x}^3_{1, t} - {x}_{1,t} - {x}_{2,t}
      \end{bmatrix} + w_t,
\end{equation*}
where $w_t\sim\mathcal{N}(\cdotx\vert 0,0.01I_2)$.
The data sampled from this black-box system is used as input for \lucid.
Given
\begin{align*}
       & \X = [ -3, 2.5 ] \times [ -2 , 1 ]                                                 \\
       & \X_0 = [ 1 , 2 ] \times [ -0.7 , 0.3 ] \cup [ -1.8 , -1.4 ] \times [ -0.1 , 0.1 ]  \\
       & \qquad \cup [-1.4, -1.2] \times [-0.5 , 0.1]                                       \\
       & \X_U = [ 0.4 , 0.6 ] \times [ 0.2 , 0.6 ] \cup [ 0.6 , 0.7 ] \times [ 0.2 , 0.4 ],
\end{align*}
we want to ensure that the system, starting in $\X_0$, does not enter the unsafe regions $\X_U$ within $T=5$ time steps.
The system dynamics and safety specification can be visualized in Figure~\ref{fig:model-barrier3}.
The kernel \hp were set to $\sigma_f=1$, $\sigma_l=[2.99, 4.63]$, and $\lambda=10^{-8}$.
The complete configuration for the \barrIII benchmark is shown in Listing~\ref{lst:barrier3}, while the barrier function is plotted in Figure~\ref{fig:CSBarr3}.
A list of experiments using different combinations of frequencies and lattice sizes, showing their impact on performance and final result, is presented in Table~\ref{tab:results-barrier3}.
\lstinputlisting[language=yaml,style={bgnonumbers},caption={Configuration for \barrIII.},captionpos=b,label={lst:barrier3}]{code/barrier3.yaml}

\begin{figure}[ht]
      \centering
      \input{figures/model-barrier3-noise.pgf}
      \caption{Visualization of the stochastic system behavior (based on 10 random transitions from an $100\times 100$ grid of initial states) and safety specification for the \barrIII benchmark.}
      \label{fig:model-barrier3}
\end{figure}

\begin{table}[tb]
      \centering
      \begin{tabular}{cccccccc}
            \toprule
            \textbf{Freq.} & \textbf{Lattice} & $\eta$ & $\gamma$ & $c$  & \textbf{Runtime} & \textbf{Runtime Mac} & \textbf{Safety} \\
                           & \textbf{Size}    &        &          &      & [mm:ss]          & [mm:ss]              & \textbf{Prob.}  \\ % Units
            \midrule
            15             & $800^2$          & 0.55   & 2        & 0.11 & 66:56            &                      & 42.25\%         \\
            16             & $700^2$          & 0.53   & 2        & 0.12 & 52:51            &                      & 41.98\%         \\
            14             & $800^2$          & 0.58   & 2        & 0.11 & 45:44            &                      & 41.66\%         \\
            13             & $800^2$          & 0.57   & 2        & 0.12 & 35:51            &                      & 40.72\%         \\
            14             & $700^2$          & 0.59   & 2        & 0.11 & 32:42            &                      & 40.57\%         \\
            13             & $700^2$          & 0.60   & 2        & 0.12 & 25:40            &                      & 39.69\%         \\
            15             & $600^2$          & 0.59   & 2        & 0.12 & 25:12            &                      & 39.57\%         \\
            12             & $800^2$          & 0.61   & 2        & 0.11 & 27:38            &                      & 39.40\%         \\
            14             & $600^2$          & 0.60   & 2        & 0.12 & 20:29            &                      & 39.24\%         \\
            13             & $600^2$          & 0.61   & 2        & 0.12 & 16:39            &                      & 38.50\%         \\
            12             & $700^2$          & 0.63   & 2        & 0.12 & 19:10            &                      & 38.43\%         \\
            12             & $600^2$          & 0.64   & 2        & 0.12 & 11:53            &                      & 37.38\%         \\
            15             & $450^2$          & 0.63   & 2        & 0.12 & 12:07            &                      & 36.15\%         \\
            14             & $450^2$          & 0.63   & 2        & 0.12 & 9:58             &                      & 36.07\%         \\
            14             & $400^2$          & 0.65   & 2        & 0.13 & 7:50             &                      & 34.77\%         \\
            14             & $370^2$          & 0.66   & 2        & 0.13 & 6:49             & 149:48               & 33.58\%         \\
            16             & $370^2$          & 0.64   & 2        & 0.13 & 11:35            &                      & 33.55\%         \\
            12             & $370^2$          & 0.71   & 2        & 0.12 & 4:38             &                      & 32.57\%         \\
            10             & $600^2$          & 0.73   & 2        & 0.12 & 46:40            &                      & 32.46\%         \\
            9              & $800^2$          & 0.74   & 2        & 0.12 & 53:07            &                      & 31.45\%         \\
            10             & $500^2$          & 0.76   & 2        & 0.12 & 45:03            &                      & 31.07\%         \\
            9              & $700^2$          & 0.75   & 2        & 0.12 & 56:43            &                      & 30.78\%         \\
            11             & $350^2$          & 0.75   & 2        & 0.12 & 3:14             &                      & 30.07\%         \\
            9              & $600^2$          & 0.77   & 2        & 0.12 & 33:57            &                      & 29.91\%         \\
            10             & $400^2$          & 0.79   & 2        & 0.12 & 34:09            &                      & 29.37\%         \\
            9              & $500^2$          & 0.79   & 2        & 0.12 & 33:40            &                      & 28.74\%         \\
            10             & $350^2$          & 0.81   & 2        & 0.12 & 2:34             &                      & 27.92\%         \\
            9              & $400^2$          & 0.80   & 2        & 0.12 & 13:07            &                      & 27.29\%         \\
            10             & $300^2$          & 0.84   & 2        & 0.12 & 8:18             &                      & 26.26\%         \\
            9              & $350^2$          & 0.83   & 2        & 0.12 & 2:02             &                      & 25.96\%         \\
            9              & $300^2$          & 0.85   & 2        & 0.13 & 6:35             & 1:21                 & 24.54\%         \\
            \bottomrule
      \end{tabular}
      \caption{\barrIII results, sorted by the last column. For each combination of number of frequencies $M$ and lattice size (i.e., number of lattice points per dimension), we report the values of $c$, $\gamma$, $\lambda$, the runtime, and the achieved lower bound on the safety probability.}
      \label{tab:results-barrier3}
\end{table}

\subsection{Overtaking}

We consider a scenario where an autonomous vehicle controlled by a \gls{nn} is overtaking another vehicle.
The dynamics of the ego vehicle are given by Dubin's car model with additive noise $w = \smash{\begin{bmatrix}w_t^1 & w_t^2 & w_t^3\end{bmatrix}\T}$ where each component $w_t^i$ is drawn from a zero-mean Gaussian with standard deviation $0.01$, $0.01$, and $0.001$, respectively.
The steering wheel angle is supplied by the \gls{nn} controller and we travel at a fixed velocity.
Given $N=1000$ sample transitions and the sets
\begin{align*}
       & \X = [ 1, 90 ] \times [ -7, 19 ] \times [ -\pi, \pi ],
       &                                                        & \X_0 = [ 1, 3 ] \times [ -1, 1 ] \times [ -0.5, 0.5 ],
       &                                                        & \X_U = [ 1, 90 ] \times [ -7, -6 ] \times [ -\pi, \pi ]                                                                       \\
       &                                                        &                                                         &  &  & \qquad \cup [ 1, 90 ] \times [ 18, 19 ] \times [ -\pi, \pi ]  \\
       &                                                        &                                                         &  &  & \qquad \cup [ 40, 45 ] \times [ -6, 6 ] \times [ -\pi, \pi ],
\end{align*}
we want to ensure that the system, starting in $\X_0$, does not enter the unsafe regions $\X_U$ within $T=5$ time steps.
The sampled transitions and safety specification are visualized in Figure~\ref{fig:model-overtaking}.
The kernel \hp are set to $\sigma_f=7$, $\sigma_l=[10, 7, 5]$, and $\lambda=10^{-5}$.
The complete configuration for the \overtaking benchmark is shown in Listing~\ref{lst:overtaking}.
A list of experiments using different combinations of frequencies and lattice sizes, showing their impact on performance and final result, is presented in Table~\ref{tab:results-overtaking}.
\lstinputlisting[language=yaml,style={bgnonumbers},caption={Configuration for \overtaking. The transition samples were generated by a separate script and appended to the configuration, abbreviated here for conciseness.},captionpos=b,label={lst:overtaking}]{code/overtaking.yaml}

\begin{figure}[ht]
      \centering
      \input{figures/model-overtaking.pgf}
      \caption{Visualization of $N=200$ sample transitions and safety specification for the \overtaking benchmark.}
      \label{fig:model-overtaking}
\end{figure}

\begin{table}[tb]
      \centering
      \begin{tabular}{ccccccc}
            \toprule
            \textbf{Freq.} & \textbf{Lattice} & $\eta$ & $\gamma$ & $c$  & \textbf{Runtime} & \textbf{Safety} \\
                           & \textbf{Size}    &        &          &      & [mm:ss]          & \textbf{Prob.}  \\ % Units
            \midrule
            5              & $99^3$           & 0.09   & 0.5      & 0.05 & 46:33            & 30.31\%         \\
            5              & $100^3$          & 0.10   & 0.5      & 0.05 & 49:21            & 29.39\%         \\
            5              & $88^3$           & 0.09   & 0.5      & 0.05 & 25:52            & 28.73\%         \\
            5              & $77^3$           & 0.10   & 0.5      & 0.05 & 13:49            & 24.94\%         \\
            5              & $66^3$           & 0.12   & 0.5      & 0.05 & 8:16             & 16.61\%         \\
            4              & $90^3$           & 0.10   & 0.5      & 0.06 & 11:21            & 11.79\%         \\
            4              & $81^3$           & 0.11   & 0.5      & 0.06 & 8:08             & 9.52\%          \\
            5              & $55^3$           & 0.13   & 0.5      & 0.06 & 4:15             & 9.21\%          \\
            4              & $72^3$           & 0.10   & 0.5      & 0.06 & 5:10             & 8.46\%          \\
            5              & $44^3$           & 0.16   & 0.5      & 0.06 & 2:06             & 4.84\%          \\
            4              & $63^3$           & 0.12   & 0.5      & 0.07 & 3:25             & 4.19\%          \\
            \bottomrule
      \end{tabular}
      \caption{\overtaking results, sorted by the last column. For each combination of number of frequencies $M$ and lattice size (i.e., number of lattice points per dimension), we report the values of $c$, $\gamma$, $\lambda$, the runtime, and the achieved lower bound on the safety probability.}
      \label{tab:results-overtaking}
\end{table}

\begin{figure}[ht]
      \centering
      \input{figures/contour-overtaking.pgf}
      \caption{Contour plot of the lower bound on the safety probability with respect to the number of frequencies $M$ and the lattice size per dimension for the \overtaking benchmark.}
      \label{fig:contour-overtaking}
\end{figure}